---
title: "Predicting Self-Rated Mental Health Based on Demographic and Family Traits"
author: "James Bao, Alan Chen, Xinyi Zhang, Rose"
date: "10/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract
Predicting mental health in the context of the nuclear family. 
Does having children have a positive (fulfilled) or negative impact (stressed)? 
Does being married have a positive or negative impact?

# Introduction

# Data
The dataset we used in our modeling is the 2017 General Social Survey (Family cycle). The following sections will discuss how the data was collected, what the key features of the dataset are, and what the data looks like.

## Data Collection
From February 1, 2017 to November 30, 2017, Statistics Canada gathered data on the Canadian family unit by conducting voluntary telephone interviews. Their target population was all non-institutionalized individuals living in Canada, aged 15 or older. Cross-sectional sampling was conducted in a two-stage design. The stratified simple random sampling method was used in the first stage. The stage one sampling frame consisted of the telephone numbers of Canadian households obtained from the Census and Statistic Canada's dwelling dataset. Stratifying the sample frame by geographic location, the sampled population consisted of chosen households in each of the 10 provinces.

In the second stage, a single family member 15 or older was randomly chosen to take the survey. in the second stage, the sampling frame is the . The sample is the chosen

-methodology: how they found people, population, frame, sample, what sampling approach,

trade-offs, 
non-response, cost
-what is good and bad about the questionnaire

## Data Characteristics
-key features, strengths, weaknesses
-discuss variables: similar that you don't use, construct variables by combining various ones

## Data Visualization
-what does the data look like
-plot the raw data

# Model

Bayes' Theorem for Naive Bayes Classifier:
$P(c|x)=\frac{P(x|c)P(c)}{P(x)}$


```{r eval=F}
# install.packages('tidyverse')
# install.packages("sjstats")
# install.packages("ROCR")
# install.packages("brms")
# install.packages("modelr")
install.packages("nnet")
# install.packages("tidybayes")
require(nnet)
library(tidyverse) # for data manipulation and plots
```
```{r}

# using 12 rows of dummy data until we get the real data

data <- tibble(
  sex = c("M", "F", "M", "M", "F", "F", "F" ,"M", "F", "M", "F", "M"),
  martial_status = c("Married", "Divorced", "Single, never married", "Living common-law", "Widowed", "Separated", "Married", "Divorced", "Single, never married", "Married", "Divorced", "Single, never married"),
  age = c("under 18", "18-34", "35-50", "50-70", "over 70", "under 18", "18-34", "35-50", "under 18", "18-34", "35-50", "18-34"),
  education = c("Bachelor's", "High school", "College", "Trade Cert", "University degree above Bachelor's level", "University degree below Bachelor's level", "Bachelor's", "High school", "College", "Trade Cert", "Bachelor's", "High school"),
  has_regilious_afflication = c("Y", "Y", "N", "Y", "Y", "N", "Y", "Y", "N", "Y", "Y", "N"),
  has_children = c("Y", "N", "Y", "N", "Y", "N", "Y", "N", "Y", "N", "N", "N"),
  selfrated_mental_health = c("Fair", "Good", "Very Good", "Excellent", "Poor", "Fair", "Good", "Very Good", "Good", "Very Good", "Excellent", "Good")
)
data


# add 1 after all x-y combinations to avoid zero frequency problem

# count y
poor <- data %>% filter(selfrated_mental_health=="Poor") %>% tally() + 1
fair <- data %>% filter(selfrated_mental_health=="Fair") %>% tally() + 1
good <- data %>% filter(selfrated_mental_health=="Good") %>% tally() + 1
vgood <- data %>% filter(selfrated_mental_health=="Very Good") %>% tally() + 1
excellent <- data %>% filter(selfrated_mental_health=="Excellent") %>% tally() + 1

total_mental <- poor + fair + good + vgood + excellent

# count sex given y
male_poor <- data %>% filter(sex=="M" & selfrated_mental_health=="Poor") %>% tally() + 1
male_fair <- data %>% filter(sex=="M" & selfrated_mental_health=="Fair") %>% tally() + 1
male_good <- data %>% filter(sex=="M" & selfrated_mental_health=="Good") %>% tally() + 1
male_vgood <- data %>% filter(sex=="M" & selfrated_mental_health=="Very Good") %>% tally() + 1
male_excellent <- data %>% filter(sex=="M" & selfrated_mental_health=="Excellent") %>% tally() + 1
female_poor <- data %>% filter(sex=="F" & selfrated_mental_health=="Poor") %>% tally() + 1
female_fair <- data %>% filter(sex=="F" & selfrated_mental_health=="Fair") %>% tally() + 1
female_good <- data %>% filter(sex=="F" & selfrated_mental_health=="Good") %>% tally() + 1
female_vgood <- data %>% filter(sex=="F" & selfrated_mental_health=="Very Good") %>% tally() + 1
female_excellent <- data %>% filter(sex=="F" & selfrated_mental_health=="Excellent") %>% tally() + 1

total_male <- male_poor + male_fair + male_good + male_vgood + male_excellent
total_female <- female_poor + female_fair + female_good + female_vgood + female_excellent

# count has_children given y
has_children_poor <- data %>% filter(has_children=="Y" & selfrated_mental_health=="Poor") %>% tally() + 1
has_children_fair <- data %>% filter(has_children=="Y" & selfrated_mental_health=="Fair") %>% tally() + 1
has_children_good <- data %>% filter(has_children=="Y" & selfrated_mental_health=="Good") %>% tally() + 1
has_children_vgood <- data %>% filter(has_children=="Y" & selfrated_mental_health=="Very Good") %>% tally() + 1
has_children_excellent <- data %>% filter(has_children=="Y" & selfrated_mental_health=="Excellent") %>% tally() + 1
no_children_poor <- data %>% filter(has_children=="N" & selfrated_mental_health=="Poor") %>% tally() + 1
no_children_fair <- data %>% filter(has_children=="N" & selfrated_mental_health=="Fair") %>% tally() + 1
no_children_good <- data %>% filter(has_children=="N" & selfrated_mental_health=="Good") %>% tally() + 1
no_children_vgood <- data %>% filter(has_children=="N" & selfrated_mental_health=="Very Good") %>% tally() + 1
no_children_excellent <- data %>% filter(has_children=="N" & selfrated_mental_health=="Excellent") %>% tally() + 1

total_has_children <- has_children_poor + has_children_fair + has_children_good + has_children_vgood + has_children_excellent
total_no_children <- no_children_poor + no_children_fair + no_children_good + no_children_vgood + no_children_excellent

# repeat for the rest of the independent var



## example: predict mental health state given male, has children
# calculating prob(poor)
class_prior_prob_poor <- poor / total_mental
likelihood_poor <- (male_poor/poor) * (has_children_poor/poor)
predictor_prior_prob_poor <- (total_male/total_mental) * (total_has_children/total_mental)
prob_poor <- likelihood_poor * class_prior_prob_poor / predictor_prior_prob_poor

prob_poor

# repeat this process to get probability for fair, good, vgood, excellent
# the classification will be the class with the largest probability

# note that the probability value themselves should not be taken seriously since naive bayes is a bad estimator (so we cannot say how confident we are in this classification)





```


```{r}
poll <- as_tibble(data.frame(read_csv("gss_cleaned.csv")))

poll <- poll %>% select(age, sex, marital_status, self_rated_mental_health)

poll <- head(poll, 1000)
poll
```
```{r}
# poll <- mutate(poll, )
# model <- brm(formula = self_rated_mental_health ~ age+sex+marital_status,
#              data = poll,
#              family = multinomial(),
#              seed = 69)

model <- multinom(self_rated_mental_health ~ age + sex + marital_status, 
                  data = poll)
summary(model)

```

```{r}
head(fitted(model))

input <- data.frame(age = c(21.5), sex = c("Male"), marital_status = c("Single, never married"))
predict(model, newdata = input, "probs")
```





# Results
# Discussion
## Weaknesses
## Next Steps

# References
Interview method: https://www.statcan.gc.ca/eng/survey/household/4501
Detailed information about GSS 2017: https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&Id=335816
Questionnaire: https://www23.statcan.gc.ca/imdb/p3Instr.pl?Function=assembleInstr&lang=en&Item_Id=335815#qb345205
